---
title: "HMWK2"
author: "Jiawen"
date: "2022-10-06"
output: html_document
---

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(lubridate)
library(psych)
library(ggplot2)
library(data.table)
library(dplyr)
library(leaflet)
```

```{r read data}
chsindividual <- fread ("chs_individual.csv")
chsregional   <- fread ("chs_regional.csv")
```


```{r}
#merge two datasets using the location variable
chsmerged <-
  chsindividual %>%
  full_join(chsregional, by="townname")
```

#1.After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches. have question here
```{r}
dim(chsmerged)
dim(chsindividual)
```
The two datasets have same number of rows which is 12, so it seems that there is no duplicates

#In the case of missing values, impute data using the average within the variables “male” and “hispanic.”
```{r}
# mutate male and hispanic, then calculate the mean of new value? BMI FEV--lecture 5
which(is.na(chsindividual$male))
which(is.na(chsindividual$hispanic))

chsmerged [, bmiimp := fcoalesce(bmi, mean(bmi, na.rm = TRUE)),
by = .(male,hispanic)]

chsmerged [, fevimp := fcoalesce(fev, mean(fev, na.rm = TRUE)),
by = .(male,hispanic)]

  #dat[, temp_imp := fcoalesce(temp, mean(temp, na.rm = TRUE)),
    #by = .(STATE, year, month, day)]
```
it shows there is no missing value in case of Male and hispanic, we input the missing values of FEV, BMI by male and hispanic

#2.1 create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24)
```{r}
    chsmerged[, obesity_level := fifelse(bmiimp<14, "underweight", 
    fifelse(bmiimp<22, "normal",
    fifelse(bmiimp<=24, "overweight", "obese")))]
```

#2.2 To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r}
   chsmerged[, .(
  N          = .N,
  minbmi     = min(bmiimp),
  maxbmi     = max(bmiimp)
  ), by = obesity_level][order(obesity_level)]
```

#3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.
```{r}
    chsmerged [, smoke_gas_exposure := fifelse(
    smoke == 0 & gasstove == 0, "noexposure",
    fifelse(smoke == 1 & gasstove == 0, "smokeexposure",
    fifelse(smoke == 0 & gasstove == 1, "gasexposure","both")))]
```

#4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r}
##by town
chsmerged %>%
  group_by(townname) %>%
  summarise(
    avgfev = mean(fevimp, na.rm=T),
    sdfev = sd(fevimp, na.rm=T),
    propasthma = mean(asthma, na.rm=T)
  ) %>%
  knitr::kable()
```

```{r}
##by sex
chsmerged %>%
  group_by(male) %>%
  summarise(
    avgfev = mean(fevimp, na.rm=T),
    sdfev = sd(fevimp, na.rm=T),
    propasthma = mean(asthma, na.rm=T)
  ) %>%
  knitr::kable()
```

```{r}
##by obesity level
chsmerged %>%
  group_by(obesity_level) %>%
  summarise(
    avgfev = mean(fevimp, na.rm=T),
    sdfev = sd(fevimp, na.rm=T),
    propasthma = mean(asthma, na.rm=T)
  ) %>%
  knitr::kable()
```

```{r}
##by smoke_gas_exposure
chsmerged %>%
  group_by(smoke_gas_exposure) %>%
  summarise(
    avgfev = mean(fevimp, na.rm=T),
    sdfev = sd(fevimp, na.rm=T),
    propasthma = mean(asthma, na.rm=T)
  ) %>%
  knitr::kable()
```

#EDA 1.Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
#1. What is the association between BMI and FEV (forced expiratory volume)? 
#2. What is the association between smoke and gas exposure and FEV? 
#3. What is the association between PM2.5 exposure and FEV?
```{r}
  chsmerged %>%
  ggplot(aes(x = bmiimp, y = fevimp, color=townname, line=townname)) +
  geom_point() +
  geom_smooth(se = F, method = "lm")
  lm(fevimp ~ bmiimp, data = chsmerged)
```
 bmi and fev are positively related 
 
#EDA 2.Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
chsmerged %>%
  ggplot(aes(x=fevimp,fill=obesity_level))+
  geom_histogram(aes(x=fevimp,fill=obesity_level))+
  scale_fill_brewer(palette='PRGn')+
  labs(title = "Histogram of FEV by BMI category", x="forced expiratory volume")

chsmerged %>%
  ggplot(aes(x=fevimp,fill=smoke_gas_exposure))+
  geom_histogram(aes(x=fevimp,fill=smoke_gas_exposure))+
  scale_fill_brewer(palette='PRGn')+
  labs(title = "Histogram of FEV by smoke/gas exposure", x="forced expiratory volume")
```
 
The amount of People whose BMI is "normal" is greater than other categories, and from this graph we can say its relationship with FEV it is basically near to normally distribution 

The amount of people who expose to gas is significantly higher than other gas or smoke exposure categories, and from this graph we can say its relationship with FEV it is basically near to normally distribution 

#EDA 3. Barchart of BMI by smoke/gas exposure.
```{r}
chsmerged %>%
  ggplot(mapping = aes(x = obesity_level, fill = smoke_gas_exposure))+
  geom_bar() +
    labs(title = "Bar Graph of BMI by Smoke/Gas Exposure", x = "BMI")  
```
No matter what type of BMI that people belong to, most of them are proposed to gas. 
#EDA 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}
chsmerged %>%
  ggplot() + 
    stat_summary(mapping = aes(x = obesity_level, y = fevimp),
    fun.min = min,
    fun.max = max,
    fun = median)+
    labs(title = "Statistical summary graphs of FEV by BMI", x = "obesity level", y = "forced expiratory volume")
```
The mean of FEV in different obesity level vary significantly. People who are obese tend to have higher FEV value

```{r}
chsmerged %>%
  ggplot() + 
    stat_summary(mapping = aes(x = smoke_gas_exposure, y = fevimp),
    fun.min = min,
    fun.max = max,
    fun = median)+
    labs(title = "Statistical summary graphs of FEV by smoke/gas exposure category", x = "smoke/gas exposure", y = "forced expiratory volume")
```
the mean of FEV in different smoke/gas exposure groups are approximate 

#EDA 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}
# Generating a color palette
#pm25.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=chsmerged$pm25_mass)
#chsmap <- leaflet(chsmerged) %>%
  #addProviderTiles('CartoDB.Positron') %>% 
  #addCircles(lat=~lat, lng=~lon,
     #label = ~`pm25_mass`, color = ~ pm25.pal(pm25_mass),
     #opacity = 1, fillOpacity = 1, radius = 60
    #) %>%
  # And a pretty legend
  #addLegend('bottomleft', pal=pm25.pal, values=chsmerged$pm25_mass,
          #title='PM2.5 mass', opacity=1)
```
shows error but do not know the reason why
#EDA 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}


```


