---
title: "LAB10"
author: "Jiawen"
date: "2022-11-02"
output: 
    github_document:
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN="http://cran.rstudio.com"))
```

```{r}
# install.packages(c("RSQLite", "DBI"))
if(!require(RSQLite)) install.packages("RSQLite")
library(RSQLite)
library(DBI)
```

```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```

```{sql, connection=con}
PRAGMA table_info(actor)
```

# exercise 1 
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

SELECT
FROM
ORDER by
```{r}
dbGetQuery(con,"
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
lIMIT 5
 ")
```

#try in SQL directly
```{sql, connection=con}
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
```

#exercise 2 
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.
```{r}
dbGetQuery(con,"
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
ORDER BY last_name
 ")
```

#Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{r}
dbGetQuery(con,"
SELECT DISTINCT customer_id
FROM rental
WHERE date(rental_date) = '2005-07-05'
")
```

```{r}
dbGetQuery(con,"
SELECT customer_id,
count(*) AS N
FROM rental
WHERE date(rental_date) = '2005-07-05'
GROUP BY customer_id
")
```

#Exercise 4.1
Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r}
dbGetQuery(con,"
SELECT *
FROM payment
WHERE amount IN(1.99, 7.99, 9.99)
")
```

#exercise Exercise 4.2
Construct a query that retrives all rows from the payment table where the amount is greater then 5
```{r}
dbGetQuery(con,"
SELECT *
FROM payment
WHERE amount >5 AND amount < 8
LIMIT 10
")
```

#Exercise 5
Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.
```{r}
dbGetQuery(con,"
PRAGMA table_info(customer)
")
```

```{r}
dbGetQuery(con,"
SELECT c.customer_id, c.last_name, p.payment_id, p.amount
FROM customer AS c INNER JOIN payment AS p
  ON c.customer_id = p.customer_id
WHERE c.last_name == 'DAVIS'
/*where c.last_name == 'DAVIS' */
")
```

#Exercise 6.1
Use COUNT(*) to count the number of rows in rental
```{r}
dbGetQuery(con,"
SELECT COUNT(*) AS count
FROM rental
")
```

#Exercise 6.2
Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id
```{r}
dbGetQuery(con,"
SELECT customer_id, COUNT(*) AS count
FROM rental
GROUP BY customer_id
LIMIT 8
")
```

#Exercise 6.3
Repeat the previous query and sort by the count in descending order
```{r}
dbGetQuery(con,"
SELECT customer_id,
COUNT(*) AS count
FROM rental
GROUP BY customer_id
ORDER BY count DESC
LIMIT 8
")
```

#Exercise 6.4
Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{r}
dbGetQuery(con,"
SELECT customer_id,
COUNT(*) AS count
FROM rental
GROUP BY customer_id
ORDER BY count DESC
LIMIT 8
")
```

#Exercise 7.1
Modify the above query to do those calculations for each customer_id
```{r}
dbGetQuery(con,"
SELECT MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       SUM(amount) AS sumpayment,
       AVG(amount) AS avgpayment
FROM payment
")
```

#Exercise 7.2
do those calculations for each customers 
Modify the above query to only keep the customer_ids that have more then 5 payments
```{r}
dbGetQuery(con,"
SELECT customer_id,
       COUNT(*) AS N,
       MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       SUM(amount) AS sumpayment,
       AVG(amount) AS avgpayment
FROM payment
GROUP BY customer_id
HAVING N>5
")
```

#Cleanup
Run the following chunk to disconnect from the connection.
```{r}
dbDisconnect(con)
```
