
install.packages("tidyverse")
install.packages("dplyr")
install.packages("data.table")
installed.packages("ggplot2")
library(ggplot2)
library(tidyverse)
library(data.table)
library(dplyr)

#1. Read in the data using data.table(). For each of the two datasets, check the dimensions, headers, footers, variable names and variable types. Check for any data issues, particularly in the key variable we are analyzing. Make sure you write up a summary of all of your findings.
```{r}
data2004 <- data.table::fread("2004.csv")
data2019 <- data.table::fread("2019.csv")

#check the dimensions
dim(data2004)
dim(data2019)

#check the heraders
head(data2004)
head(data2019)

#check the footers
tail(data2004)
tail(data2019)

#check the variable names and types
str(data2004)
str(data2019)

#check the summary
summary(data2004)
summary(data2019)

#especially check the key variable, which is daily concentrations of PM2.5
summary(data2004$'Daily Mean PM2.5 Concentration')
summary(data2019$'Daily Mean PM2.5 Concentration')

#It has 20 variables for each dataset. The mean number of PM2.5 was 13.13 in 2004, 7.739 in 2019. All the values decrease from in 2019 compared with those in 2004
```

#2. Combine the two years of data into one data frame. 
#Use the Date variable to create a new column for year, which will serve as an identifier. 
#Change the names of the key variables so that they are easier to refer to in your code.
```{r}
library(dplyr)
  total <- rbind (data2004, data2019) 
  data2004$Year <- "2004"
  data2019$Year <- "2019"
  total <- rbind(data2004, data2019)
  total <- rename(total, c("PM2.5" = "Daily Mean PM2.5 Concentration", "Lat"="SITE_LATITUDE", "Long"="SITE_LONGITUDE"))
```

#3. Create a basic map in leaflet() that shows the locations of the sites (make sure to use different colors for each year). Summarize the spatial distribution of the monitoring sites.
```{r}
library(ggplot2)
ggplot(map_data("state"), aes(x = long, y = lat)) +
  geom_map(aes(map_id = region), map = map_data("state"), col = "lightgrey", fill = "gray") +
  geom_point(
    data    = total,
    mapping = aes(x = Long, y = Lat, col = Year),
    inherit.aes = FALSE, cex = 0.1, alpha = 0.5
    )
#We can tell from the map that the number of sites in 2019 significantly increases than in 2004
```
#4.Check for any missing or implausible values of PM in the combined dataset. 
#Explore the proportions of each and provide a summary of any temporal patterns you see in these observations.
```{r}
#total number and proportion of missing values
is.na(total)
which(is.na(total))
sum(is.na(total)/prod(dim(total)))*100
sum(is.na(total$`PM2.5`/prod(dim(total$`PM2.5`))))*100
sum(total$PM2.5<0)
#total proportion of missing vale is 0.391%, no missing PM data. However, there is 283 negative PM2.5 value, which is not sensible. Therefore, I will remove them before moving to the next step
```

#5. Explore the main question of interest at three different spatial levels. 
#Create exploratory plots (e.g. boxplots, histograms, line plots) and summary statistics that best suit each level of data. Be sure to write up explanations of what you observe in these data.
#state
#boxlot
```{r}
library(ggplot2)
totalafter[STATE == "California"] %>%
ggplot()+
geom_boxplot(mapping = aes(x = Year, y = PM2.5, group = Year))
```
# At State level, the overall number of PM2.5 in 2019 is lower than 2004

#county
#lineplots
```{r}
library(ggplot2)
ggplot(totalafter)+ 
 geom_line(mapping=aes(x=PM2.5, y=COUNTY))
```
#Mariposa has singificant higher PM2.5 than other counties
