---
title: "HW4"
author: "Jiawen"
date: "2022-11-16"
output: 
       github_document: 
             html_preview: false
       html_document: default
       pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem 1: Make sure your code is nice
Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google
```{r}
if(!require(matrixStats)) install.packages("matrixStats") 
library(matrixStats)
set.seed(2315)
dat <- matrix(rnorm(4 * 3), nrow = 4)

fun1 <- function(mat) { #mat is a matrix 
  n <- nrow(mat)
  ans <- double(n) #ans is a name of variable
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  # YOUR CODE HERE
rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  # YOUR CODE HERE
rowCumsums(mat)
  }


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat),unit = "nanoseconds"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "nanoseconds"
)
```

#Problem 2: Make things run faster with parallel computing
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

#In order to get accurate estimates, we can run this function multiple times, with the following code:
```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

#Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream(): refer to slide 30
```{r}
# YOUR CODE HERE
library(parallel)

system.time({
# This runs the simulation a 4,000 times, each with 10,000 points
  cl <- makePSOCKcluster(4)  #making cluster, and its name is cl
  clusterSetRNGStream(cl, 123)  #setting the sead
  
  ans <- unlist(parLapply(cl, 1:4000, sim_pi, n = 10000))
  parallel::stopCluster(cl)
  print(mean(ans))
})

```

#SQL
##Setup a temporary database by running the following chunk
```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
dbListTables(con)
```
###view table 
```{sql, connection=con}
      PRAGMA table_info(film_category)
```

```{sql, connection=con}
      PRAGMA table_info(film)
```

```{sql, connection=con}
      PRAGMA table_info(category)
```

###Q1.How many many movies is there avaliable in each rating catagory.
```{r}
dbGetQuery(con,"
SELECT  rating, COUNT(*) AS count
FROM  film
GROUP BY rating
")
```

###Q2.What is the average replacement cost and rental rate for each rating category.
```{r}
dbGetQuery(con,"
SELECT rating,
       AVG (replacement_cost) AS avgreplacementcost
FROM film
GROUP BY rating
")

dbGetQuery(con,"
SELECT rating,
       AVG (rental_rate) AS avgrentalrate
FROM film
GROUP BY rating
")
```

###Q3. Use table film_category together with film to find the how many films there are witth each category ID
```{sql, connection=con}
SELECT film_category.category_id, COUNT(*) AS count
FROM film
LEFT JOIN film_category ON film_category.film_id = film.film_id
GROUP BY film_category.category_id;
```
###Q4. Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql, connection=con}
SELECT film_category.category_id, category.name, COUNT(*) AS count
FROM film
LEFT JOIN film_category ON film_category.film_id = film.film_id
LEFT JOIN category ON film_category.category_id = category.category_id
GROUP BY film_category.category_id
ORDER BY count DESC;
```

```{r}
# clean up
dbDisconnect(con)
```
