---
title: "README.md"
author: "Jiawen"
date: "2022-09-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN="http://cran.rstudio.com"))
```


```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")
```

```{r}
dim(met)
```
```{r}
head(met)
```

```{r}
tail(met)
```

```{r}
str(met)
```

```{r}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```

```{r}
summary(met$temp)
```

```{r}
summary(met$elev)

summary(met$wind.sp)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)

met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)

met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)

elev <- met[elev==max(elev)]
summary(elev)

cor(elev$temp, elev$wind.sp, use="complete")

cor(elev$temp, elev$hour, use="complete")

cor(elev$wind.sp, elev$day, use="complete")

cor(elev$wind.sp, elev$hour, use="complete")

cor(elev$temp, elev$day, use="complete")
```

install.packages("leaflet")

```{r}
hist(met$elev, breaks=100)

hist(met$temp)

hist(met$wind.sp)

library(tidyverse)

#leaflet(elev) %>%
#  addProviderTiles('OpenStreetMap') %>% 
#  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')

plot(elev$date, elev$wind.sp, type='l')
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
```             



##2. Prepare the data Remove temperatures less than -17C Make sure there are no missing data in the key variables coded as 9999, 999, etc
```{r}
library(data.table)
met<- met[temp>-17][elev == 9999, elev := NA]
```

Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
```{r create-ymd, cache=TRUE}
met <- met[ ,ymd :=as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.
# the year of 
```{r}
met[, table(week(ymd))]
met <- met[week(ymd) == 31]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
```{r check-mean}
met_avg <- met[, .(
  temp = mean(temp,na.rm=T),
  rh = mean(rh,rm=T),
  wind.sp = mean(wind.sp,rm=T),
  vis.dist = mean(vis.dist,rm=T),
  dew.point = mean(dew.point,rm=T),
  lat = mean(lat,rm=T),
  lon = mean(lon,rm=T),
  elev = mean(elev,rm=T)
),by="USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}
met_avg[,region := fifelse(lon >= -98 & lat > 39.71,"NE",
               fifelse(lon < -98 & lat > 39.71,"NW",
               fifelse(lon < -98 & lat <= 39.71,"SW","SE")))]
table(met_avg$region)
```
Create a categorical variable for elevation as in the lecture slides
```{r}
met_avg[, elev := fifelse(elev > 252,"high","low")]
```

##3.Use geom_violin to examine the wind speed and dew point temperature by region
```{r}
met_avg[!is.na(region)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x=1, y=dew.point, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```
##4.Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region
```{r scatterplot-dewpoint-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp) & !is.na(dew.point)] %>% 
  ggplot(mapping = aes(x=wind.sp, y=dew.point)) + 
  geom_point(mapping=aes(color=region)) + 
  geom_smooth(mapping = aes(linetype = region))+
  facet_wrap(~region, nrow = 2)
```
comment on these results 

##5.Use geom_bar to create barplots of the weather stations by elevation category coloured by region

##6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
```{r}
 met_avg[!is.na(region)& !is.na(dew.point)] %>%
  ggplot() + 
    stat_summary(mapping = aes(x = region, y = dew.point),
    fun.min = min,
    fun.max = max,
    fun.data = mean_sdl)
```
##7.Make a map showing the spatial trend in relative h in the US
# Generating a color palette
```{r}
#install.packages("terra")
#install.packages("raster")
#install.packages("leaflet")
#library("leaflet")
#rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'),domain=met_avg2$rh)
#leaflet()%>%
#addTiles() %>%
#the looks of the map
#addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  #addCircles(
    #lat = ~lat, lng=~lon,
  # HERE IS OUR PAL!
    #label = ~paste0(round(rh), ' C'), color = ~ rh.pal(rh),
    #opacity = 1, fillOpacity = 1, radius = 500
   # ) %>%

  # And a pretty legend
 # addLegend('bottomleft', pal=rh.pal, values=met_avg2$rh,
         # title='Temperature, C', opacity=1)
```
#

